Lifting CDCL to Relational Domains

### 2016-10-14 domains.tex
ssa fig, domain edits

\section{Program and Abstract Domains}
In this section, we define a notion of program analysis constraints, concrete
domain and abstract domain. \\
\Omit {
\subsection{Programs:}
%
A {\em program} $\mathcal{P}$ is formally defined as follows.
\[
\begin{array}[t]{@{}lll}
\mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
\end{array}
\]
Consider sets of expressions $Exp$ and Boolean expressions $BExp$
over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
can take numeric values in $Val$.  A procedure is denoted as 
$\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
is an assignment, conditional, sequential concatenation or a loop. \\

\textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
where $Loc$ is the set of locations with an unique start 
location $(init)$ and error location $(err)$, $E$ is the set 
of control-flow edges that are labelled with the set 
$lbl \in Statement$.  For the purpose of illustration, we 
assume that all the procedures are inlined.  
}
\textbf{Program Analysis Constraints:}  A program $\mathcal{P}$ is represented 
as set of equations called, {\em program analysis constraints}.  The 
program analysis constraints, as shown below, is an assignment, 
or boolean expressions $BExp$ over program variables $Var$ or
a non-deterministic choice (ternary experssion) between statements $s_1$ and
$s_2$.  All the bounded loops are completely unrolled and represented using 
following constraints. Similarly, an assertion is of type $BExp$.  
\[c::= x:= exp | BExp | (exp) ? s_1, s_2\]
For example, figure~\ref{swssa} presents the program analysis constraints for 
a C program in Static Single Assignment (SSA) form.

\pscmt{we should give the (simplified) SSA for an example here}

\subsection{Concrete Semantics}
\pscmt{not sure that adds anything for TACAS}
The {\em concrete} domain is a lattice 
of concrete environments, $Env = Var \rightarrow Val$, and is 
defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
A transformer, $post_{stmt}$, of concrete domain defines 
the effect of a statement $stmt$ on the concrete domain, 
$post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.

A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
$0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.

\subsection{Template Polyhedra Abstract Domain}

\pscmt{TODO}

\paragraph{Interval Abstract Domain} \pscmt{TODO: simplify}
The {\em Interval} abstract domain is a lattice 
$ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
$ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
the set of intervals of type $[l,u]$ over numeric data 
type with $l \leq u$. The least element is $\bot$ and the 
greatest element is $\top$ which maps all variables to their
minimum ($min$) and maximum ($max$) values.  An interval 
$\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
partial order $\sle_{I}$ over elements in the set $Itv$ is 
given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
The galois connection between the concrete environment and intervals is 
given as follows.
\[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
\[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
For a program with $N$ variables, there are a total of 
$2*N$ intervals. The intervals restrict the set of values 
taken by program variables at each program point, starting 
from $[-\infty, \infty]$.

\paragraph{Octagon Abstract Domain} \pscmt{TODO: simplify}
The {\em Octagon} abstract domain is a lattice 
$OctDom = (OctElm, \sle_{O}, \join_{O}, \meet_{O})$, where
$OctElm: (Var \times Var \rightarrow (\mathbb{R} \cup \{\infty\})) \cup \bot$. 
The least element is $\bot$ that contains all unsatisfiable 
set of inequalities and the greatest element is $\top$ which 
maps the bounds of all octagonal inequalities to $\infty$. 
The partial order $\sle_{O}$ over elements in the set $OctElm$ is 
given by $O_1 \sle_{O} O_2$ iff the bounds of each inequalities in $0_1$ 
is included (by $\leq$ order) in the bounds of corresponding inequalities 
in $O_2$, that is, the octagons are ordered by the inclusion relations.

The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
and is computed by taking piece-wise maximum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
computed by taking piece-wise minimum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
the octagons that abstract the same concrete values.

The octagon domain is a relational abstract domain that permits $2n^2$ 
linear inequalities between $n$ program variables.  The octagonal 
inequalities are of two types: binary or unary inequalities as shown below.
\[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
An element in octagon domain, $OctElm$, is a conjunction of such 
inequalities and is called an octagon.  Figure~\ref{octagon} shows an 
example octagon and its associated inequalities.   
%
\begin{wrapfigure}{r}{5.5cm}
\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
\end{wrapfigure} 
%
For building program analyzers using octagon domains, the domain also 
provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
The widening operator is used to accelerate convergence for loops in the program
and has a quadratic complexity in the number of variables.  If the bound of an 
octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
to $\infty$.  However, the closure operator is often used to reduce the degree 
of over-approximation resulting from the join operation. Closure always produces 
unique octagonal inequalities, but it is very expensive to compute because of 
cubic complexity in the number of program variables. An example of closure
operation is shown below.
\[((x-y \leq 4) \wedge (y-z \leq 5))^* \implies ((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]

>>>>

\section{Preliminaries}
This section explains the program model and the abstract domains that
we are going to use.
% \subsection{Programs:}
% %
% A {\em program} $\mathcal{P}$ is formally defined as follows.
% \[
% \begin{array}[t]{@{}lll}
% \mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
% \mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
% \mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
% \mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
% \end{array}
% \]
% Consider sets of expressions $Exp$ and Boolean expressions $BExp$
% over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
% can take numeric values in $Val$.  A procedure is denoted as 
% $\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
% is an assignment, conditional, sequential concatenation or a loop. \\
% \textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
% where $Loc$ is the set of locations with an unique start 
% location $(init)$ and error location $(err)$, $E$ is the set 
% of control-flow edges that are labelled with the set 
% $lbl \in Statement$.  For the purpose of illustration, we 
% assume that all the procedures are inlined.

\begin{figure}
\[
\begin{array}{rcl}
g_0 &=& \true \\
\multicolumn{3}{l}{0 \leq v\leq N} \\
g_1 &=& g_0\wedge c\\
x_0 &=& v \\
x_1 &=& -v \\
x_2 &=& g_1?x_0:x_1 \\
g_2 &=& g_1 \vee g_0\wedge \neg c\\
z &=& x_2 * x_2 \\
\multicolumn{3}{l}{z<0 \wedge g_2}
\end{array}
\]
\caption{\label{fig:ssa}
SSA for the example in Figure~\ref{fig:example}
}
\end{figure}

%-------------------------------------------------------------------------------
\subsection{Program Representation}  
%
We consider loop-free programs (respectively bounded unwindings of
programs with loops or recursion) with safety properties given as
assertions in the code.
%
Such a program is represented as set $\mathcal{P}$ of
\emph{statements} $s$.  These are constraints as typically obtained by
translating the program into single static assignment (SSA) form via a
data flow analysis.
%
Figure~\ref{fig:ssa} shows the constraints for the example program in
Figure~\ref{fig:example}.
%
Assignments such as \texttt{x=v;} become equalities $x_1=v$ where the
left-hand side variable gets a fresh name (subscript).
%
Control flow is encoded into guard variables such as $g_1=g_0\wedge c$.
%
Data flow joins become conditional expressions, e.g.\ $x_3=g_1?x_1:x_2$.
%
Assertions $\mathcal{A}$ are constraints $a$ of the form $g_3
\Rightarrow z\geq 0$, meaning that if $g_3$ holds (i.e.\ the assertion
is reachable) then the assertion must hold.

%For example, figure~\ref{swssa} presents the program analysis constraints for 
%a C program in Static Single Assignment (SSA) form.

% \subsection{Concrete Semantics}
% \pscmt{not sure that adds anything for TACAS}
% The {\em concrete} domain is a lattice 
% of concrete environments, $Env = Var \rightarrow Val$, and is 
% defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
% A transformer, $post_{stmt}$, of concrete domain defines 
% the effect of a statement $stmt$ on the concrete domain, 
% $post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.  
% 
% A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
% where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
% of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
% $0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
% if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.

%-------------------------------------------------------------------------------
\paragraph{Safety problem}  
%
Based on above program representation, the safety problem that our
ACDL solver will solve is the following:
%
\begin{equation}\label{eq:safety}
\bigwedge_{s\in\mathcal{P}} s \wedge \neg \bigwedge_{a\in\mathcal{A}} a
\text{ is unsatisfiable if and only if the program is safe.}
\end{equation}

%-------------------------------------------------------------------------------
\subsection{Abstract Domain}
%-------------------------------------------------------------------------------
The results that we present on lifting ACDL to relational domains hold
for any relational domain.
% 
In our prototype implementation and for illustrative purposes in this
paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
$\mathcal{B}$ is the Boolean domain
$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$,
$b$ the number of Boolean variables in the program, and $\mathcal{R}$
is a \emph{relational} domain over the numerical (bitvector) variables
in their respective data types (currently, signed and unsigned
integers are supported).
%
We instantiate $\mathcal{R}$ with the template polyhedra domain.

For notational convenience we will denote elements of
$\mathcal{B}^b$ by their concretisations to propositional formulae,
e.g.\ for the Boolean variable vector $(g_1,c)$ with abstract value
$(\false,\top)$ we will write $\neg g_1$.

%-------------------------------------------------------------------------------
\paragraph{Template Polyhedra Abstract Domain}
%
The template polyhedra domain~\cite{sriram} represents sets $X$ of
valuations of a vector of variables $\vec{x}$ such that
$\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
by a special element to denote the minimal element $\bot$ of the lattice.
%
The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is
applied component-wise. There are several optimisation-based
techniques for computing the domain
operations~\cite{sriram,GS07b,BJKS15}. In our implementation we use
the strategy iteration approach available in 2LS~\cite{BJKS15}.
%
For notational convenience we will use conjunctions of linear
inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
of the domain. $\true$ is $\top$ and $\false$ is $\bot$.

% \paragraph{Interval Abstract Domain} 
% The {\em Interval} abstract domain is a lattice 
% $ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
% $ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
% the set of intervals of type $[l,u]$ over numeric data 
% type with $l \leq u$. The least element is $\bot$ and the 
% greatest element is $\top$ which maps all variables to their
% minimum ($min$) and maximum ($max$) values.  An interval 
% $\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
% partial order $\sle_{I}$ over elements in the set $Itv$ is 
% given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
% A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
% A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
% $\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
% The galois connection between the concrete environment and intervals is 
% given as follows.
% \[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
% \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
Template polyhedra subsume, for example, interval and octagon domains.

\begin{wrapfigure}{r}{4.5cm}
\vspace*{-3ex}
\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
\caption{An example of Octagon}\label{octagon}
\vspace*{-3ex}
\end{wrapfigure}

For a program with $N$ variables, the matrix $\mat{C}$ for intervals
generates at most $2N$ inequality constraints, one for the upper and lower
bounds of each variable.
%
For octagons, we have at most $2N^2$ inequalities, one for the upper
and lower bounds of each variable and sums and differences for each
pair of variables. Figure~\ref{octagon} shows an 
example octagon and its associated inequalities.

The interval domain is a non-relational domain because a single
inequality only contains a single variable.
%
The octagon domain, however, is relational.  A fundamental difference
between relational and non-relational domains is that relational
domains require the computation of a \emph{closure} in order to obtain
a normal form that is usually required for efficient domain operations.
An example of a closure computation is shown below.
\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
The closure computes all implied domain constraints.  For octagons,
the closure operation is very expensive (cubic complexity in the
number of program variables).
%
We will see that our algorithm performs closure computations lazily,
postponing them to the point where they are actually necessary.

% \paragraph{Octagon Abstract Domain} \pscmt{TODO: simplify}
% The {\em Octagon} abstract domain is a lattice 
% $OctDom = (OctElm, \sle_{O}, \join_{O}, \meet_{O})$, where
% $OctElm: (Var \times Var \rightarrow (\mathbb{R} \cup \{\infty\})) \cup \bot$. 
% The least element is $\bot$ that contains all unsatisfiable 
% set of inequalities and the greatest element is $\top$ which 
% maps the bounds of all octagonal inequalities to $\infty$. 
% The partial order $\sle_{O}$ over elements in the set $OctElm$ is 
% given by $O_1 \sle_{O} O_2$ iff the bounds of each inequalities in $0_1$ 
% is included (by $\leq$ order) in the bounds of corresponding inequalities 
% in $O_2$, that is, the octagons are ordered by the inclusion relations.

% The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
% (v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
% and is computed by taking piece-wise maximum of bounds of corresponding 
% octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
% However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
% (v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
% computed by taking piece-wise minimum of bounds of corresponding 
% octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
% octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
% the octagons that abstract the same concrete values.

% The octagon domain is a relational abstract domain that permits $2n^2$ 
% linear inequalities between $n$ program variables.  The octagonal 
% inequalities are of two types: binary or unary inequalities as shown below.
% \[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
% An element in octagon domain, $OctElm$, is a conjunction of such 
% inequalities and is called an octagon.  
%
% For building program analyzers using octagon domains, the domain also 
% provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
% The widening operator is used to accelerate convergence for loops in the program
% and has a quadratic complexity in the number of variables.  If the bound of an 
% octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
% to $\infty$.  However, the closure operator is often used to reduce the degree 
% of over-approximation resulting from the join operation.

\pscmt{TODO from here}

### 2016-10-14 example.tex
ssa fig, domain edits

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void main() {
 int x,z;
 int val;
 assume(val>=0 && val<=N);
 _Bool c;
 if(c)
   x = val;
 else 
   x = -val;
 z = x * x;
 assert(z>=0);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{\label{example}
C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
\pscmt{change \&\& in the CFG to $\wedge$; the constraint towards to error does not show up correctly; in the second oct analysis box put $\wedge$ between the constraints; add an arrow from the interval Error box with three dots to indicate that the analysis continues (even though you explain it in the text); shorten ``val'' to ``v''}
}
\end{figure}
Let us consider a simple program in Figure~\ref{example} that performs 
multiplication of two machine integers and checks that the result of the 
multiplication is always positive for an integer input $x$.  To avoid 
overflow error, we assume $x$ has an upper bound $N$.  The control 
flow graph is shown in the middle column of Figure~\ref{example}.

We analyze the program using three different analyzers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
\Omit {
The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
the solver performs a restart, presumably to recover from bad decisions.  
}
Standard BMC tools convert the program to a bit-vector equation and pass 
them to a SAT solver.  However, standard forward interval analysis is 
too imprecise to verify the safety of the above program.  The 
imprecision is due to control-flow join at node $n4$.  State-of-the-art 
abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
by manually annotating the code with partition directives at $n1$, 
to prove safety.  The partition directive helps the tool to analyze 
each program execution paths separately.  To do so, Astr{\'e}e 
requires a product of trace partitioning domain and an interval domain 
to determine safety.  However, a relational domain such as octagons is 
able to prove the program without any partition.  In general, the 
imprecision is either intended by the tool because such high 
precision analysis is normally not required for runtime error 
analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.  
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
\hline
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{solver}
\end{table}
%
The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
octagon domain (top) and an interval domain (bottom).  The analysis 
associates an abstract element for each control location and variable, 
as shown in the abstract conflict graph in Firgure~\ref{example}.  
Octagonal constraints deduced by ACDCL solver are sufficient to prove 
safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
an automatic interval analysis is also sufficient to prove safety with 
only a single decision.  Interval constraints generated from forward 
analysis in the initial deduction phase infers that 
$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL can not prove safety.

Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
ACDCL implements several decision heuritics.  In this case, the 
branching variable $c$ is chosen by an {\em ordered} decision 
heuristics that select variables appearing in conditional branches 
first before choosing other program variables. 
The decision constrains the interval of $x:=[0,5]$.  The deductions made 
during fixed point iteration are represented by abstract conflict 
graph in right-side of figure~\ref{example}.  Nodes in the abstract 
conflict graph denotes the corresponding updates to the program 
variables in the CFG.  Interval analysis concludes that the $(ERROR:\bot)$, that 
is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
the program is {\em safe} for $c:=[1,1]$.

>>>>

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void main() {
 int x,z;
 int val;
 assume(val>=0 && val<=N);
 _Bool c;
 if(c)
   x = val;
 else 
   x = -val;
 z = x * x;
 assert(z>=0);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{\label{fig:example}
C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
\pscmt{change \&\& in the CFG to $\wedge$; the constraint towards to error does not show up correctly; in the second oct analysis box put $\wedge$ between the constraints; add an arrow from the interval Error box with three dots to indicate that the analysis continues (even though you explain it in the text); shorten ``val'' to ``v''}
}
\end{figure}
Let us consider a simple program in Figure~\ref{fig:example} that performs 
multiplication of two machine integers and checks that the result of the 
multiplication is always positive for an integer input $x$.  To avoid 
overflow error, we assume $x$ has an upper bound $N$.  The control 
flow graph is shown in the middle column of Figure~\ref{fig:example}.

We analyze the program using three different analyzers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
\Omit {
The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
the solver performs a restart, presumably to recover from bad decisions.  
}
Standard BMC tools convert the program to a bit-vector equation and pass 
them to a SAT solver.  However, standard forward interval analysis is 
too imprecise to verify the safety of the above program.  The 
imprecision is due to control-flow join at node $n4$.  State-of-the-art 
abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
by manually annotating the code with partition directives at $n1$, 
to prove safety.  The partition directive helps the tool to analyze 
each program execution paths separately.  To do so, Astr{\'e}e 
requires a product of trace partitioning domain and an interval domain 
to determine safety.  However, a relational domain such as octagons is 
able to prove the program without any partition.  In general, the 
imprecision is either intended by the tool because such high 
precision analysis is normally not required for runtime error 
analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.  
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
\hline
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{solver}
\end{table}
%
The right-side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
octagon domain (top) and an interval domain (bottom).  The analysis 
associates an abstract element for each control location and variable, 
as shown in the abstract conflict graph in Firgure~\ref{fig:example}.  
Octagonal constraints deduced by ACDCL solver are sufficient to prove 
safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
an automatic interval analysis is also sufficient to prove safety with 
only a single decision.  Interval constraints generated from forward 
analysis in the initial deduction phase infers that 
$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL can not prove safety.

Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
ACDCL implements several decision heuritics.  In this case, the 
branching variable $c$ is chosen by an {\em ordered} decision 
heuristics that select variables appearing in conditional branches 
first before choosing other program variables. 
The decision constrains the interval of $x:=[0,5]$.  The deductions made 
during fixed point iteration are represented by abstract conflict 
graph in right-side of figure~\ref{fig:example}.  Nodes in the abstract 
conflict graph denotes the corresponding updates to the program 
variables in the CFG.  Interval analysis concludes that the $(ERROR:\bot)$, that 
is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
the program is {\em safe} for $c:=[1,1]$.
